\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-12-04 resubmit_mnras/response.txt

Some of he figures (e.g., Fig. 4) are too small for comfortable reading, while others (e.g., Fig. 6) are unnecessarily large. In Fig. 9, the axes need to be explained and labelled (at least the time range); I'm not sure why you include broken bolometers as "typical" - the space could be better used! In figures with many panels, it helps to label the panels (a), (b), etc. and refer to the panels by label - see the instructions to authors.

- Fig.4 had been modified, using a smaller number of sub-windows and a
  larger font it is now more readable.

- Fig.6 has been made smaller.

- Fig.9 is now Fig. 10, and now has axis labels and an improved
  caption - however, we elected to leave in the "broken" bolometers as
  this is a fairly typical portion of the focal plane (remembering
  that there is a ~70% yield); this gives the reader a good visual
  sense of that.

@@

- done

Assistant Editor's Comments:

Holland et al and Dempsey et al are not yet 'in press' as they have not been accepted. Please cite these as 'submitted to MNRAS'.

- in the References the "in press" has been replaced with "submitted"
  for these two papers

### 2012-12-04 smurf_paper.tex
minor changes based on comments from co-authors

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers (with a useful yield of
  69\%) in each of two simultaneous imaging bands centred over 450 and
  850\,\micron. The camera is operated by scanning across the sky and
  recording data at a rate of 200\,Hz. As the largest of a new
  generation of multiplexed kilopixel bolometer cameras operating in
  the (sub)millimetre, \scuba\ data analysis represents a significant
  challenge.  We describe the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers, with
  execution times that are typically shorter than the observing times.
  SMURF is used in an automated setting, both at the telescope for
  real-time feedback to observers, as well as for the production of
  science products for the JCMT Science Archive at the Canadian
  Astronomy Data Centre. Three detailed case studies are used to: (i)
  explore convergence properties of the map-maker using simple prior
  constraints (Uranus -- a point source); (ii) achieve the white-noise
  limit for faint point-source studies (extragalactic blank-field
  survey of the Lockman Hole); and (iii) demonstrate that our strategy
  is capable of recovering angular scales comparable to the size of
  the array footprint (approximately 5\,arcmin) for bright extended
  sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $10,240$ nominal detectors \citep[of which $\sim$69\%
work and are typically useful,][]{holland2012}, \scuba\ is presently
the largest of a new generation of multiplexed kilopixel
(sub)millimetre bolometer cameras, which also includes the cameras for
the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
experiments for studying anisotropies in the Cosmic Microwave
Background (CMB) using similar technology, the latter of which uses the
same time-domain multiplexed readout electronics as \scuba\
\citep{battistelli2008}.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales commensurate with the observing
times (or significantly faster), on single high-end desktop
computers. For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline offering feedback to
observers at the telescope. The pipeline is also used to generate
products for the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomy Data Centre (CADC). Details of the
pipeline design are given in \citet{gibb2005} and
\citet{2008AN....329..295C}.

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (though not located in
the focal plane itself, it is certainly correlated with the
temperature of the detectors), the line-of-sight atmoshperic opacity
as derived from the JCMT water vapour monitor (WVM), and the telescope
pointing, for two data sets, before and after the upgrades that
followed S2SRO.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{skynoise}
\caption{Illustration of the dominance of atmospheric variations on
  long-timescales using 35\,min. of data from the fully-commissioned
  instrument, observation 18 on 20110324. Red and blue lines show the
  850 and 450\,micron\ common-mode signals, corresponding to the left-
  and right- vertical axes, respectively. The black line shows the
  line-of-sight opacity from the WVM as a proxy for atmospheric
  emission (arbitrary scale). Bolometer baseline drifts on timescales
  $\gsim$100\,s are clearly correlated with the atmosphere in this
  case.}
\label{fig:skynoise}
\end{figure}

The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} are
dominated at low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise,
Rather than underlying variations in the atmosphere, or ``sky
noise''. However, only $\sim$100\,s of data were analysed. In
Fig.~\ref{fig:skynoise}, an order-of-magnitude longer
(fully-commissioned) data set is plotted from observation 18 on
20110324. The common-mode signals from the 450 and 850\,\micron\ focal
planes are compared with the (arbitrarily scaled) line-of-sight
opacity at 183\,GHz from the WVM. Unlike the earlier plots, it is
clear that the bolometer baseline drifts on timescales $\gsim$100\,s
are tightly correlated with the atmospheric opacity (especially at
850\,\micron), which serves as a proxy for atmospheric emission. The
fact that the correlation between the WVM and the 450\,micron\
bolometers is weaker either reflects an underlying difference in the
atmosphere at higher frequencies, or the presence of some other
significant low-frequency noise source.

While the relative importance of sky noise at low-frequencies is
highly time- and scan pattern-dependent, in general it does not have a
major impact on map-making. Upon common-mode removal (which is an
integral part of our map-making strategy) uncorrelated noise remains
(e.g., solid lines in Fig~\ref{fig:pspec}). As we will see in
Section~\ref{sec:pca} and \ref{sec:experimental}, this remaining noise
does not have a smoothly-varying correlation pattern across the focal
plane one might expect if it were due to resolved ``clouds'' of
atmospheric emission. A high-pass filter above $\sim$0.1--1.0\,Hz is
usually employed remove this residual noise
(Section~\ref{sec:components}), which will also remove any sky noise
that is not already subtracted by the common-mode.

@@

In general, SMURF is highly configurable, including many options for
both the pre-processing stage and the iterative model components
(which models are used, what order they are calculated in, how they
are calculated).  In Sections~\ref{sec:dataprep} and
\ref{sec:components} we describe the data pre-processing steps and
iterative algorithm in detail. Section~\ref{sec:converge} explores
convergence tests and degeneracies between model components. Finally,
Section~\ref{sec:performance} summarizes the mapping performance of
\scuba\ using SMURF.

@@

Finally, data that are taken while the telescope is either moving
extremely slowly, or extremely fast, are flagged and ignored when
estimating the map. When the telescope is moving slowly, astronomical
signals appear at low frequencies in the bolometer data that are
dominated by $1/f$ noise (in the extreme case of no motion, the
bolometers have zero sensitivity). When the telescope is moving
significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
response of the bolometers, and the anti-aliasing filter in the
readout electronics (Section~\ref{sec:bolos}) can significantly
distort signals.  The default allowable range of speed is
30--1000\,arcsec\,s$^{-1}$, which primarily rejects short periods when
the telescope was stationary prior to commencing the scan pattern
(generally speaking the speed is constant once the scan begins),
although the scan speed can briefly exceed the upper limit, even for a
requested speed of 600\,arcsec, if the pointing system is having
trouble keeping-up for large-area scans at high elevation.

@@

For this reason, the usual practice is to estimate a single
\model{COM} model by examining data from all four sub-arrays in each
waveband, since this allows spatial structure on the scale of the
whole focal plane to be recovered. However, sometimes there is
evidence that the common-mode differs from one array to another, and
so an option exists to estimate a separate \model{COM} model for each
individual subarray, with a consequent lowering in the scale of
spatial structure that can be recovered.

@@

From a scientific perspective, the most important goal of map-making
is to efficiently use all of the available data to achieve the
greatest sensitivity that the instrument is capable of. Generally
speaking, only a small percentage of the data are flagged as unusable
during map-making beyond the $\sim$69\% detector yield.  As an
example, for the 450\,\micron\ scan of Uranus that will be presented
in Section~\ref{sec:point}, during pre-processing: 31.00\% of the data
(1587/5120 bolometers) are flagged due to data acquisition problems
(dead or deactivated bolometers, or non-linear flatfields); 0.12\% of
the data are flagged due to steps (including a small region about the
steps, Section~\ref{sec:steps}); and 0.83\% if the data are flagged
because the telescope was stationary
(Section~\ref{sec:flagbadbol}). Once the iterative solution begins: 55
spikes are detected using the map-based despiker ($1.2\times10^{-6}$
\% of the data, Section~\ref{sec:ast}); and 0.72\% of the data are
flagged as outliers compared to the common-mode
(Section~\ref{sec:comgai}).

The white noise levels of \scuba\ bolometers are measured between 2
and 10\,Hz in the PSDs (as described in Section~\ref{sec:noi}),
leading to noise equivalent powers (NEPs, with units W\,s$^{1/2}$) --
the \rms\ time-series noise in a 1\,s integration \citep[see section
3.5.2 in][]{holland2012}. The NEPs are then converted into more useful
noise equivalent flux densities (NEFDs, with units Jy\,s$^{1/2}$)
through multiplication by the extinction correction and the FCF. For
an ideal scan pattern in which every bolometer spends an equal amount
of time observing each point on the sky, the expected noise in a map
pixel is $\mathrm{NEFD_{eff}} t^{-1/2}$, where $t$ is the integration
time, and the effective NEFD, $\mathrm{NEFD_{eff}} = (1/\sum
\mathrm{NEFD}_i^{-2})^{1/2}$, is the sensitivity of the entire array
including all bolometers $i$. Experimentally it has been found that
the map noise based on the weighted scatter of the data
(Eq.~\ref{eq:varmap}), and the scatter of map pixels in small
apertures, both correlate well with this expected limiting noise
performance for maps that are larger than the array footprint (for
smaller maps the assumption of uniform coverage is incorrect). As an
example of observing under favourable conditions with line-of-sight
opacities of 0.87 and 0.34 at 450 and 850\,\micron\ \citep[zenith sky
opacities at 225\,GHz of 0.04 and 0.065 at 450 and 850\,\micron,
respectively, at an average airmass of 1.2, from table 3
in][]{holland2012}, the mapping speed is approximately
$6\times10^{-5}$ and $3\times10^{-3}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$,
at 450 and 850\micron, for the default pixel sizes of 2 and 4\,arcsec,
respectively. If maps are high-pass filtered and cross-correlated with
the PSF for the optimal extraction of point sources (see
Section~\ref{sec:cosmo}), the mapping speeds increase to approximately
$5\times10^{-4}$ and
$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$, independently of
the chosen map pixel sizes. See the on-line \scuba\ integration time
calculator for mapping speeds under arbitrary observing conditions
and scan
modes\footnote{\url{http://www.jach.hawaii.edu/jac-bin/propscuba2itc.pl}}.

@@

This example illustrates the need for constraints in the map solution
in many situations. For calibrators (and other previously known
bright, compact sources), a good, simple prior is to constrain the map
to a value of zero away from the known locations of emission (part of
the \model{AST} model calculation, Section~\ref{sec:ast}).  In
Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
explicitly to zero beyond a radius of 60\,arcsec from the location of
Uranus (much larger than the FWHM of the main lobe), for all but the
final iteration. In this case, the map converges after 6 iterations,
and the ringing has been effectively removed. The attenuation from the
previous reduction is now removed, and Uranus again has a peak value
of 0.358\,pW. For reference, the Uranus model distributed with
Starlink predicts a peak brightness of 176\,Jy on this date, yielding
an FCF of 492 Jy\,pW$^{-1}$, which is well within the typical range
for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
details]{dempsey2012}. Since the map is now flat away from the source,
and constrained to a value of zero, it is appropriate for performing
aperture photometry directly, with no need for an additional reference
annulus.

@@

First, the angular power spectrum of noise is estimated from a
``jackknife map'': maps are produced from two independent halves of
the total data set, and the jackknife signal in a map pixel,
$S_\mathrm{JK}$, and its variance, $\sigma^2_\mathrm{JK}$ are
estimated from the two input map fluxes, $S_1$, $S_2$, and the
corresponding variances, $\sigma^2_1$, and $\sigma^2_2$ as,
%
\begin{eqnarray}
S_\mathrm{JK} = \frac{S_1 - S_2}{2}, \\
\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}.
\end{eqnarray}
%
Provided that the noise in one half of the data is uncorrelated with
that from the other half, the signal in the jackknife map should
resemble noise drawn from the same parent distribution as that of the
real map. The astronomical signal, however, should be cleanly removed
(provided that there are no strong time-varying signals, and also
assuming that errors due to calibration between the two halves are
insignificant). The approach we have taken to minimize systematics is
to produce the two maps using odd and even scan numbers (i.e., each
map will contain a nearly uniformly-spaced sampling of data across the
full data set).

@@

Next, we analyse the angular power spectral densities (PSDs) of the
maps to understand the signal and noise properties of the map-maker in
the region of synthetic sources, as a function of filter scale. In
Fig.~\ref{fig:m17_def_ps}a we show the raw PSDs for the input
synthetic signal (thick black line), the output map signals (thin
solid lines), and the jackknife maps (dashed lines). Colours encode
the filter scales used: 150\,arcsec (red); 300\,arcsec (orange);
600\,arcsec (green); and 900\,arcsec (blue). Note that, with the
exception of the synthetic data, we only plot the PSDs down to the
second-lowest spatial frequency bin of
$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, due to the fact
that the smallest is very poorly sampled, and therefore noisy. As the
filter edge is increased, more power is measured in the map PSDs at
larger scales. However, much of this power is clearly produced by
noise which appears in the jackknife PSDs. We estimate the map-maker
transfer function as the ratio between the portion of the signal PSDs
not produced by noise, to the input PSD, or
%
\begin{equation}
G(k) = \frac{P_\mathrm{M}(k) - P_\mathrm{JK}(k)}{P_\mathrm{S}(k)},
\end{equation}
%
where $k$ is the spatial frequency, and the subscripts ``M'', ``JK'',
and ``S'' refer to the signal map, jackknife map, and synthetic map,
respectively.

@@

These tests are then repeated using the bright extended reduction, in
Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
reduction over the baseline reduction is that the transfer functions
fall more slowly at large angular scales, accompanied by a slower
increase in noise; in other words, there is greatly improved \snr\ at
large angular scales (an obvious conclusion given the appearance of
the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
filter edge, the map response is still above 80 per cent right out to
the largest scale accurately measured in the PSDs, 533\,arcsec, which
is about the largest scale that should be recoverable, given the size
of the array footprint and the fact that we use common-mode rejection.
Another interesting feature of these reductions is that the increase
in small-scale noise as the filter edge is increased is not as drastic
as in the baseline reduction, and the \snr\ improves at scales
$\gsim$60\,arcsec.

@@

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the construction of \scuba\ were provided by the Canada
Foundation for Innovation.  This research was supported in part by the
Natural Sciences and Engineering Research Council of Canada.  EC
thanks CANARIE/CANFAR for additional funding.  The authors thank the
members of the SCUBA-2 commissioning team for testing the map-maker
and reporting anomalies; in particular Antonio Chrysostomou and
Jessica Dempsey.  We also thank Mark Halpern, Matthew Hasselfield, and
Gaelen Marsden for many useful discussions; observers who provided
helpful feedback, especially David Nutter and Todd McKenzie; and
Mandana Amiri and Dan Bintley for consultations regarding the SCUBA-2
Multi-Channel Electronics.  We acknowledge the contributions of Dennis
Kelly, Alex van Engelen and Jennifer Balfour for early investigations
related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
Gaudet for being on the Critical Design Review panel. We thank Per
Friberg and Gary Davis for their helpful comments on the
manuscript. Finally, we thank the anonymous referee for identifying
several key areas that required clarification.

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  nominally consisting of 5120 bolometers in each of two simultaneous
  imaging bands centred over 450 and 850\,\micron. The camera is
  operated by scanning across the sky and recording data at a rate of
  200\,Hz. As the largest of a new generation of multiplexed kilopixel
  bolometer cameras operating in the (sub)millimetre, \scuba\ data
  analysis represents a significant challenge.  We describe the
  production of maps using the Sub-Millimetre User Reduction Facility
  (SMURF) in which we have adopted a fast, iterative approach to
  map-making that enables data reduction on single, modern, high-end
  desktop computers, with execution times that are typically shorter
  than the observing times.  SMURF is used in an automated setting,
  both at the telescope for real-time feedback to observers, as well
  as for the production of science products for the JCMT Science
  Archive at the Canadian Astronomy Data Centre. Three detailed case
  studies are used to: (i) explore convergence properties of the
  map-maker using simple prior constraints (Uranus -- a point source);
  (ii) achieve the white-noise limit for faint point-source studies
  (extragalactic blank-field survey of the Lockman Hole); and (iii)
  demonstrate that our strategy is capable of recovering angular
  scales comparable to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17).
\end{abstract}

@@

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: discovering
through blind surveys a class of massive dusty star-forming galaxies
in the early ($z>2$) Universe, now referred to as submillimetre
galaxies, or SMGs; the characterisation of the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; and identifying debris disks
around nearby stars, helping us understand the early stages of planet
formation.  With $10,240$ nominal detectors \citep[of which $\sim$70\%
work and are typically useful,][]{holland2012}, \scuba\ is presently
the largest of a new generation of multiplexed kilopixel
(sub)millimetre bolometer cameras, which also includes the cameras for
the South Pole Telescope \citep[SPT,][]{carlstrom2011} and the Atacama
Cosmology Telescope \citep[ACT,][]{swetz2011}, both dedicated
experiments for studying anisotropies in the Cosmic Microwave
Background (CMB) using similar technology, the latter of which uses the
same time-domain multiplexed readout electronics as \scuba\
\citep{battistelli2008}.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales commensurate with the observing
times (or significantly faster), on single high-end desktop computers
using the . For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline \citep[based on the
Observatory Reduction and Acquisition Control - Data Reduction system,
or ORAC-DR,][]{gibb2005} offering feedback to observers at the
telescope. The pipeline is also used to generate products for the JCMT
Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
Astronomy Data Centre (CADC). Details of the pipeline design are given
in \citet{gibb2005} and \citet{2008AN....329..295C}.

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (though not located in
the focal plane itself, it is certainly correlated with the
temperature of the detectors), the line-of-sight atmospheric opacity
as derived from the JCMT water vapour monitor (WVM), and the telescope
pointing, for two data sets, before and after the upgrades that
followed S2SRO.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{skynoise}
\caption{Illustration of the dominance of atmospheric variations on
  long-timescales using 35\,min. of data taken in opaque, and
  highly-variable weather conditions (line-of-sight 450\,\micron\
  opacity ranging from 12 to 22) from the fully-commissioned
  instrument, observation 18 on 20110324. Red and blue lines show the
  850 and 450\,\micron\ common-mode signals, corresponding to the left-
  and right- vertical axes, respectively. The black line shows the
  line-of-sight opacity from the WVM as a proxy for atmospheric
  emission (arbitrary scale). Bolometer baseline drifts on timescales
  $\gsim$100\,s are clearly correlated with the atmosphere in this
  case.}
\label{fig:skynoise}
\end{figure}

The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} were
taken during fairly stable weather conditions (although with different
450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
and post-upgrade data sets, respectively), and are dominated at
low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
underlying variations in the atmosphere, or ``sky noise''. However,
only $\sim$100\,s of data were analysed. In Fig.~\ref{fig:skynoise},
an order-of-magnitude longer (fully-commissioned) data set is plotted
from observation 18 on 20110324, taken during more opaque, and highly
variable weather conditions (450\,\micron\ line-of-sight opacities
ranging from 12 to 22, scaled from the WVM data which are shown as a
black line). Unlike the earlier plots, it is clear that the bolometer
baseline drifts on timescales $\gsim$100\,s are tightly correlated
with the atmospheric opacity, which serves as a proxy for atmospheric
emission. The weaker correlation between the WVM and the 450\,micron\
bolometers is probably due to the atmosphere being more transmissive
at 850\micron, therefore the emission detected at 450\,\micron\ comes
from water vapour that is closer to the telescope.

While the relative importance of sky noise at low-frequencies is
highly time- and scan pattern-dependent, in general it does not have a
major impact on map-making. Upon common-mode removal (which is an
integral part of our map-making strategy) uncorrelated noise remains
(e.g., solid lines in Fig~\ref{fig:pspec}). As we will see in
Section~\ref{sec:pca} and \ref{sec:experimental}, this remaining noise
does not have a smoothly-varying correlation pattern across the focal
plane one might expect if it were due to resolved ``clouds'' of
atmospheric emission. A high-pass filter with an edge frequency in the
range $\sim$0.1--1.0\,Hz is usually employed remove this residual
noise (Section~\ref{sec:components}), which will also remove any sky
noise that is not already subtracted by the common-mode.

@@

In general, SMURF is highly configurable, including many options for
both the pre-processing stage and the iterative model components
(which models are used, what order they are calculated in, how they
are calculated).  In Sections~\ref{sec:dataprep} and
\ref{sec:components} we describe the data pre-processing steps and
iterative algorithm in detail. Section~\ref{sec:converge} explores
convergence tests and degeneracies between model components. Finally,
Section~\ref{sec:performance} summarises the mapping performance of
\scuba\ using SMURF.

@@

Finally, data that are taken while the telescope is either moving
extremely slowly, or extremely fast, are flagged and ignored when
estimating the map. When the telescope is moving slowly, astronomical
signals appear at low frequencies in the bolometer data that are
dominated by $1/f$ noise (in the extreme case of no motion, the
bolometers have zero sensitivity). When the telescope is moving
significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
response of the bolometers, and the anti-aliasing filter in the
readout electronics (Section~\ref{sec:bolos}) can significantly
distort signals.  The default allowable range of speed is
30--1000\,arcsec\,s$^{-1}$. These thresholds primarily reject short
periods when the telescope was stationary prior to commencing the scan
pattern (generally speaking the speed is constant once the scan
begins). The scan speed can also briefly exceed the upper limit, even
for a requested speed of 600\,arcsec, if the pointing system is
struggling for large-area scans at high elevation.

@@

For this reason, the usual practice is to estimate a single
\model{COM} model by examining data from all four subarrays in each
waveband, since this allows spatial structure on the scale of the
whole focal plane to be recovered. However, sometimes there is
evidence that the common-mode differs from one array to another, and
so an option exists to estimate a separate \model{COM} model for each
individual subarray, with a consequent lowering in the scale of
spatial structure that can be recovered.

@@

From a scientific perspective, the most important goal of map-making
is to efficiently use all of the available data to achieve the
greatest sensitivity that the instrument is capable of. Generally
speaking, only a small percentage of the data are flagged as unusable
during map-making beyond the $\sim$70\% detector yield.  As an
example, for the 450\,\micron\ scan of Uranus that will be presented
in Section~\ref{sec:point}, during pre-processing: 31.00\% of the data
(1587/5120 bolometers) are flagged due to data acquisition problems
(dead or deactivated bolometers, or non-linear flatfields); 0.12\% of
the data are flagged due to steps (including a small region about the
steps, Section~\ref{sec:steps}); and 0.83\% if the data are flagged
because the telescope was stationary
(Section~\ref{sec:flagbadbol}). Once the iterative solution begins: 55
spikes are detected using the map-based despiker ($1.2\times10^{-6}$
\% of the data, Section~\ref{sec:ast}); and 0.72\% of the data are
flagged as outliers compared to the common-mode
(Section~\ref{sec:comgai}).

The white noise levels of \scuba\ bolometers are measured between 2
and 10\,Hz in the PSDs (as described in Section~\ref{sec:noi}),
leading to noise equivalent powers (NEPs, with units W\,s$^{1/2}$) --
the \rms\ time-series noise in a 1\,s integration \citep[see section
3.5.2 in][]{holland2012}. The NEPs are then converted into more useful
noise equivalent flux densities (NEFDs, with units Jy\,s$^{1/2}$)
through multiplication by the extinction correction and the FCF. For
an ideal scan pattern in which every bolometer spends an equal amount
of time observing each point on the sky, the expected noise in a map
pixel is $\mathrm{NEFD_{eff}}/t^{1/2}$, where $t$ is the integration
time, and the effective NEFD, $\mathrm{NEFD_{eff}} = (1/\sum
\mathrm{NEFD}_i^{-2})^{1/2}$, is the sensitivity of the entire array
including all bolometers $i$. Experimentally it has been found that
the map noise based on the weighted scatter of the data
(Eq.~\ref{eq:varmap}), and the scatter of map pixels in small
apertures, both correlate well with this expected limiting noise
performance for maps that are larger than the array footprint (for
smaller maps the assumption of uniform coverage is incorrect). As an
example of observing under favourable conditions with line-of-sight
opacities of 0.87 and 0.34 at 450 and 850\,\micron\ \citep[zenith sky
opacities at 225\,GHz of 0.04 and 0.065 at 450 and 850\,\micron,
respectively, at an average airmass of 1.2, from table 3
in][]{holland2012}, the mapping speed is approximately
$6\times10^{-5}$ and $3\times10^{-3}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$,
at 450 and 850\micron, for the default pixel sizes of 2 and 4\,arcsec,
respectively. If maps are filtered and cross-correlated with the PSF
for the optimal extraction of point sources (see
Section~\ref{sec:cosmo}), the mapping speeds increase to approximately
$5\times10^{-4}$ and
$1.5\times10^{-2}$\,deg$^2$\,hr$^{-1}$\,mJy$^{-2}$ in this example,
independently of the chosen map pixel sizes. See the on-line \scuba\
integration time calculator for mapping speeds under arbitrary
observing conditions and scan
modes\footnote{\url{http://www.jach.hawaii.edu/jac-bin/propscuba2itc.pl}}.

@@

This example illustrates the need for constraints in the map solution
in many situations. For calibrators (and other previously known
bright, compact sources), a good, simple prior is to constrain the map
to a value of zero away from the known locations of emission (part of
the \model{AST} model calculation, Section~\ref{sec:ast}).  In
Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
explicitly to zero beyond a radius of 60\,arcsec from the location of
Uranus (much larger than the FWHM of the main lobe), for all but the
final iteration. In this case, the map converges after 6 iterations,
and the ringing has been effectively removed. The attenuation from the
previous reduction is now removed, and Uranus again has a peak value
of 0.358\,pW. For reference, the Uranus model distributed with
Starlink predicts a peak brightness of 176\,Jy on this date, yielding
an FCF of 491 Jy\,pW$^{-1}$, which is well within the typical range
for \scuba\ of $(492 \pm 67)$\,Jy\,pW$^{-1}$ \citep[see][for
details]{dempsey2012}. Since the map is now flat away from the source,
and constrained to a value of zero, it is appropriate for performing
aperture photometry directly, with no need for an additional reference
annulus.

@@

First, the angular power spectrum of noise is estimated from a
``jackknife map'': maps are produced from two independent halves of
the total data set, and the jackknife signal in a map pixel,
$S_\mathrm{JK}$, and its variance, $\sigma^2_\mathrm{JK}$ are
estimated from the two input map fluxes, $S_1$, $S_2$, and the
corresponding variances, $\sigma^2_1$, and $\sigma^2_2$ as,
%
\begin{eqnarray}
S_\mathrm{JK} = \frac{S_1 - S_2}{2}, \\
\sigma^2_\mathrm{JK} = \frac{\sigma^2_1 + \sigma^2_2}{4}.
\end{eqnarray}
%
Provided that the noise in one half of the data is uncorrelated with
that from the other half, the signal in the jackknife map should
resemble noise drawn from the same parent distribution as that of the
real map. The astronomical signal, however, should be cleanly removed
(provided that there are no strong time-varying signals, and also
assuming that errors due to calibration between the two halves are
insignificant). The approach we have taken to minimise systematics is
to produce the two maps using odd and even scan numbers (i.e., each
map will contain a nearly uniformly-spaced sampling of data across the
full data set).

@@

Next, we analyse the angular power spectral densities (PSDs) of the
maps to understand the signal and noise properties of the map-maker in
the region of synthetic sources, as a function of filter scale. In
Fig.~\ref{fig:m17_def_ps}a we show the raw PSDs for the input
synthetic signal (thick black line), the output map signals (thin
solid lines), and the jackknife maps (dashed lines). Colours encode
the filter scales used: 150\,arcsec (red); 300\,arcsec (orange);
600\,arcsec (green); and 900\,arcsec (blue). Note that, with the
exception of the synthetic data, we only plot the PSDs down to the
second-lowest spatial frequency bin of
$1.875\times10^{-3}$\,arcsec$^{-1}$, or 533\,arcsec, since the lowest
is very poorly sampled, and therefore noisy. At small scales
($\lsim$20\,arcsec) the map and jackknife PSDs flatten. However,
unlike Fig.~\ref{fig:lockman_pspec}, these white noise levels are
slightly larger in the jackknife maps, suggesting that the differences
have not completely removed the astronomical signals. A strong
possibility is that the two half-maps that go into the differences
have not converged equally, probably due to the lack of prior
constraints.  As the filter edge is increased, more power is measured
in the map PSDs at larger scales. However, much of this power is
clearly produced by noise which appears in the jackknife PSDs. We
estimate the map-maker transfer function as the ratio between the
portion of the signal PSDs not produced by noise, to the input PSD, or
%
\begin{equation}
G(k) = \frac{P_\mathrm{M}(k) - P_\mathrm{JK}(k)}{P_\mathrm{S}(k)},
\end{equation}
%
where $k$ is the spatial frequency, and the subscripts ``M'', ``JK'',
and ``S'' refer to the signal map, jackknife map, and synthetic map,
respectively.

@@

These tests are then repeated using the bright extended reduction, in
Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
reduction over the baseline reduction is that the transfer functions
fall more slowly at large angular scales, accompanied by a slower
increase in noise; in other words, there is greatly improved \snr\ at
large angular scales (an obvious conclusion given the appearance of
the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
filter edge, the map response is still above 80 per cent right out to
the largest scale accurately measured in the PSDs, 533\,arcsec, which
is about the largest scale that should be recoverable, given the size
of the array footprint and the fact that we use common-mode rejection.
Another interesting feature of these reductions is that the increase
in small-scale noise as the filter edge is increased is not as drastic
as in the baseline reduction, and the \snr\ improves at scales
$\gsim$60\,arcsec. Finally, note that both the map and jackknife white
noise levels (at scales $\lsim$20\,arcsec) are in excellent agreement,
unlike the previous example.

@@

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the construction of \scuba\ were provided by the Canada
Foundation for Innovation.  This research was supported in part by the
Natural Sciences and Engineering Research Council of Canada.  EC
thanks CANARIE/CANFAR for additional funding.  The authors thank the
members of the SCUBA-2 commissioning team for testing the map-maker
and reporting anomalies; in particular Antonio Chrysostomou and
Jessica Dempsey.  We also thank Mark Halpern, Matthew Hasselfield, and
Gaelen Marsden for many useful discussions; observers who provided
helpful feedback, especially David Nutter and Todd McKenzie; and
Mandana Amiri and Dan Bintley for consultations regarding the SCUBA-2
Multi-Channel Electronics.  We acknowledge the contributions of Dennis
Kelly, Alex van Engelen and Jennifer Balfour for early investigations
related to SMURF; and Mark Thompson, Craig Walther and S\'{e}verin
Gaudet for being on the Critical Design Review panel. We thank Per
Friberg and Gary Davis for their helpful comments on the
manuscript. Finally, we thank the anonymous referee for a very helpful
review that helped to clarify several key areas of the paper.
